R for Clinical Study Reports and Submission

### 2021-11-17 tlf-ae-spec.Rmd
streamline paragraphs

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which participants were included in each efficacy
analysis in Section 12.2, Adverse Events (AEs).

@@

In this chapter, we illustrate how to summarize simplified specific AE
information in a study.

@@

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
adae <- read_sas("adam_data/adae.sas7bdat")
```

For illustration purposes, we only provide count in the simplified
table. The percentage of participants for each AE criteria can be
calculated as in Chapter \@ref(aesummary).

In this way, let's focus on the analysis script for two advanced
features for a table layout.

@@

In the code below, we count the number of subjects in each AE term by
SOC and treatment group, and we create a new variable `order` and set it
as `0` for these counted numbers. The variable `order` will help with
the data manipulation later.

@@

In the code below, we count the number of subjects in each AE term by
SOC, AE term, and treatment group. Here we also create a new variable
`order` and set it as `1` for these counted numbers.

@@

We prepare reporting data for AE information.

@@

We prepare reporting data for analysis population.

@@

We start to define the format of the output.

To obtain the nested layout, we use the `page_by` argument in the
`rtf_body` function. By defining `page_by="AESOC"`, `r2rtf` recognize
the variable as a group indicator.

@@

More discussion on `page_by`, `group_by` and `subline_by`
features can be found in the
[`r2rtf` package website](https://merck.github.io/r2rtf/articles/example-sublineby-pageby-groupby.html.

@@

The procedure to generate a baseline characteristics table can be
summarized as follow:

- Step 1: Read data into R, i.e., `adae` and `adsl`.
- Step 2: Count the number of subjects by SOC and treatment group
  (rows with bold text) and save into `t1`.
- Step 3: Count the number of subjects in each AE term by SOC, AE
  term, and treatment group (rows without bold text) and save into `t2`.
- Step 4: Bind `t1` and `t2` by row into `t_ae`.
- Step 5: Count the number of subjects in each arm as `t_pop`.
- Step 6: Bind `t_pop` and `t_ae` by row into `tbl_ae_spec`.
- Step 7: Format the output by `r2rtf`.

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize number of participants for each specific AE in Section 12.2, Adverse Events (AEs).

@@

In this chapter, we illustrate how to summarize simplified specific AE
information for a study.

@@

```{r}
adsl <- read_sas("data-adam/adsl.sas7bdat")
adae <- read_sas("data-adam/adae.sas7bdat")
```

For illustration purpose, we only provide counts in the simplified
table. The percentage of participants for each AE can be
calculated as shown in Chapter \@ref(aesummary).

Here, we focus on the analysis script for two advanced
features for a table layout.

@@

In the code below, we count the number of participants in each AE term by
SOC and treatment arm, and we create a new variable `order` and set it
as `0`. The variable `order` can help with
the data manipulation later.

@@

In the code below, we count the number of subjects in each AE term by
SOC, AE term, and treatment arm. Here we also create a new variable
`order` and set it as `1`.

@@

We prepare reporting data for AE information using code below:

@@

We prepare reporting data for analysis population using code below:

@@

We define the format of the output as below:

To obtain the nested layout, we use the `page_by` argument in the
`rtf_body` function. By defining `page_by="AESOC"`, `r2rtf` recognizes
the variable as a group indicator.

@@

@@

More discussion on `page_by`, `group_by` and `subline_by`
features can be found on the
[`r2rtf` package website](https://merck.github.io/r2rtf/articles/example-sublineby-pageby-groupby.html).

The procedure to generate a baseline characteristics table can be
summarized as follows:

- Step 1: Read data (i.e., `adae` and `adsl`) into R.
- Step 2: Count the number of participants by SOC and treatment arm
  (rows with bold text) and save into `t1`.
- Step 3: Count the number of participants in each AE term by SOC, AE
  term, and treatment arm (rows without bold text) and save into `t2`.
- Step 4: Bind `t1` and `t2` by row into `t_ae`.
- Step 5: Count the number of participants in each arm as `t_pop`.
- Step 6: Row-wise combine `t_pop` and `t_ae` into `tbl_ae_spec`.
- Step 7: Format the output by using `r2rtf`.

### 2021-11-17 tlf-ae-summary.Rmd

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which participants were included in each efficacy analysis in Section 12.2, Adverse Events (AEs).

@@

In this chapter, we illustrate how to summarize AE information in a study.

@@

We first summarize participants in population by treatment group.

@@

- With one or more adverse events
- With drug-related adverse events
- With serious adverse events
- With serious drug-related adverse events
- Who died"

@@

We prepare reporting-ready dataset for each AE criteria.

@@

We start to define the format of the output.

@@

The procedure to generate an AE summary table can be summarized as follow:

- Step 1: Read data into R, i.e., `adae` and `adsl`.
- Step 2: Summarize participants in population by treatment group, and we name the dataset as `t_pop`.
- Step 3: Summarize participants in population by required AE criteria of interest,
  and we name the dataset as `t_ae`.
- Step 4: Rowly bind `t_pop` and `t_ae` and format it by `r2rtf`.

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to the number of participants were included in each safety analysis in Section 12.2, Adverse Events (AEs).

@@

In this chapter, we illustrate how to summarize AEs information for a study.

@@

We first summarize participants in population by treatment arm.

@@

- With one or more adverse events
- With drug-related adverse events
- With serious adverse events
- With serious drug-related adverse events
- Who died

@@

We prepare reporting-ready dataset for each AE group.

@@

We define the format of the output using code below:

@@

The procedure to generate an AE summary table can be summarized as follows:

- Step 1: Read data (i.e., `adae` and `adsl`) into R.
- Step 2: Summarize participants in population by treatment arm, and name the dataset as `t_pop`.
- Step 3: Summarize participants in population by required AE criteria of interest,
  and name the dataset as `t_ae`.
- Step 4: Row-wise combine `t_pop` and `t_ae` and format it by using `r2rtf`.

### 2021-11-17 tlf-overview.Rmd

In clinical trials, a critical step is to submit trial results to regulatory agencies.
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document)
has become a worldwide regulatory submission standard format.
For example, the United States Food and Drug Administration (US FDA) required
new drug applications and biologics license applications
[must be submitted using the eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd).
The Clinical Data Interchange Standards Consortium (CDISC) provided a
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

@@

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical analysis, a large number of tables, listings, and figures are
incorporated into the main text and appendices.
In the CDISC pilot project, an
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf)
is also provided. If you are interested in more examples of clinical study reports,
the European Medicines Agency (EMA) publishes clinical data, including
clinical study reports submitted by pharmaceutical companies
on the [EMA clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

In this part, we mainly use the R packages below to illustrate
how to deliver TLFs in CSR.

@@

`r2rtf` is an R package to create production-ready tables and figures in RTF format.
The R package is designed to

@@

We provide a brief introduction of `r2rtf` and showed how to transfer
data frames into table, listing, and figures (TLFs).

@@

To explore the basic RTF generation verbs in `r2rtf`,
we will use the dataset `r2rtf_adae` saved in the `r2rtf` package.
This dataset contains adverse events (AE) information from a clinical trial.

>>>>

In clinical trials, a critical step is to submit trial results to regulatory agencies.
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document)
has become a worldwide regulatory submission standard format.
For example, the United States Food and Drug Administration (US FDA) requires
new drug applications and biologics license applications
[must be submitted using the eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd).
The Clinical Data Interchange Standards Consortium (CDISC) provides a
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

@@

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical analysis, a large number of tables, listings, and figures are
incorporated into the main text and appendices.
In the CDISC pilot project, an
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf)
is also provided. If you are interested in more examples of clinical study reports,
you can go to the [European Medicines Agency (EMA) clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

@@

In this part, we mainly use the R packages below to illustrate
how to deliver TLFs in a CSR.

@@

`r2rtf` is an R package to create production-ready tables and figures in RTF format.
This R package is designed to

@@

We provide a brief introduction of `r2rtf` and showe how to transfer
data frames into table, listing, and figures (TLFs).

@@

To explore the basic RTF generation verbs in `r2rtf`,
we will use the dataset `r2rtf_adae` saved in the `r2rtf` package.
This dataset contains adverse events (AEs) information from a clinical trial.
